Rice yield improvements through plant breeding are offset by inherent yield declines over time

### 2018-03-11 output/breed_paper_revised.Rnw
After large yank.

Not accounting for yield decline over time, the estimated yield
improvement of \Sexpr{only_release_coef} kg ha\tsups{-1}
year\tsups{-1} (Fig. \ref{fig:wo_erosion}) is lower than the 37 to 60
kg ha\tsups{-1} year\tsups{-1} estimated by~\citet{Omar2014} in
side-by-side comparisons. To evaluate if 37 to 60 kg ha\tsups{-1}
year\tsups{-1} is reasonable, the potential yields at
the beginning and end of the study period can be compared. ~\citet{vanIttersum2013} defined potential
yield as the yield of optimum management and well-adapted cultivars given
current environmental limitations. Under this definition, potential yield
could be reasonably estimated from the yields of 'M-202' and 'M-206'
in experimental plots at the RES in the years 1984 and 2015,
respectively. Given these criteria, it is clear that potential yield
has changed little over the study period
(Fig. \ref{fig:first_yrs}). When accounting for yield decline over
time across all cultivars, there is no conclusive evidence of
increased genetic yield potential in medium grain rice cultivars
released by CCRRF's breeding improvement program over the period
\Sexpr{min(mod_data[,'year'])} to \Sexpr{max(mod_data[,'year'])}.
Grain yield over this period has stayed relatively constant at an
average \Sexpr{round(mean(mod_data[,"yield_kg"])/1000, 1)} Mg
ha\tsups{-1}. Since the estimated yield improvement spans from
\Sexpr{rel_CI[1]} to \Sexpr{rel_CI[2]} kg ha\tsups{-1} year\tsups{-1}
(Fig. \ref{fig:RES_year}; Table \ref{tbl:coef}), both increases and
decreases in the genetic yield potential over the study period are
plausible. This analysis suggests that the CCRRF breeding program has
maintained but not substantially improved the yield potential in the
face of changing biotic and abiotic pressures. This conclusion holds
true even when evaluating the cultivars which have been breed for
broad adaption and high yield potential (M-202, M-206 and
M-105). Yields of these cultivars at time of release have all yielded
between
\Sexpr{tmp=mod_data[mod_data[,"yrs_in_trial"]<=0&mod_data[,"id"]%in%c("M202","M206","M105"),];round(min(by(tmp,tmp[,"id"],function(x)mean(x[,"yield_kg"]))/1000),2)}
    and
    \Sexpr{round(max(by(tmp,tmp[,"id"],function(x)mean(x[,"yield_kg"]))/1000),2)}
    Mg ha\tsups{-1} with no indication of yields increasing
    (Fig. \ref{fig:first_yrs}).

>>>>

This analysis suggests that the CCRRF breeding program has
maintained but not substantially improved the yield potential in the
face of changing biotic and abiotic pressures over the 33 year time
frame of this study. This conclusion holds
true even when evaluating the cultivars which have been breed for
broad adaption and high yield potential (M-202, M-206 and
M-105). Yields of these cultivars at time of release have all yielded
between
\Sexpr{tmp=mod_data[mod_data[,"yrs_in_trial"]<=0&mod_data[,"id"]%in%c("M202","M206","M105"),];round(min(by(tmp,tmp[,"id"],function(x)mean(x[,"yield_kg"]))/1000),2)}
    and
    \Sexpr{round(max(by(tmp,tmp[,"id"],function(x)mean(x[,"yield_kg"]))/1000),2)}
    Mg ha\tsups{-1} with no indication of yields increasing
    (Fig. \ref{fig:first_yrs}).
